---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
#MNIST example
##Testing notebook
```{r}
library(keras)
mnist<-dataset_mnist()
str(mnist)
firstTrain<-mnist$train$x[1,,]
plot(as.raster(firstTrain,max=255))
```

##Creating input data
```{r}
train_images <- mnist$train$x
train_labels <- mnist$train$y
test_images <- mnist$test$x
test_labels <- mnist$test$y
```

##Simple NN
###Building simple neural network model
```{r}
network <- keras_model_sequential() %>% 
  layer_dense(units = 512, activation = "relu", input_shape = c(28 * 28)) %>% 
  layer_dense(units = 10, activation = "softmax")
```
###Compiling
```{r}
network %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)
```

###Reshaping and normalizing data for simple neural network. Each image is a vector of pixels, labels are categorical
```{r}
train_images <- array_reshape(train_images, c(60000, 28 * 28))
train_images <- train_images / 255

test_images <- array_reshape(test_images, c(10000, 28 * 28))
test_images <- test_images / 255

train_labels <- to_categorical(train_labels)
test_labels <- to_categorical(test_labels)
```

###Training the network
```{r, echo=TRUE, results='hide'}
network %>% fit(train_images, train_labels, epochs = 5, batch_size = 128)
```
###Accuracy metrics for the test images
```{r}
metrics <- network %>% evaluate(test_images, test_labels, verbose = 0)
metrics
```

##Convnets
###Model - convolutional layers
```{r}
model <- keras_model_sequential() %>% 
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(28, 28, 1)) %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu")

summary(model)
```

###Model - densly connected classifier network
```{r}
model <- model %>% 
  layer_flatten() %>% 
  layer_dense(units = 64, activation = "relu") %>% 
  layer_dense(units = 10, activation = "softmax")

summary(model)
```

###Reshaping data, compiling, fitting and testing
```{r, echo=TRUE, results='hide'}
mnist <- dataset_mnist()
c(c(train_images, train_labels), c(test_images, test_labels)) %<-% mnist

train_images <- array_reshape(train_images, c(60000, 28, 28, 1))
train_images <- train_images / 255

test_images <- array_reshape(test_images, c(10000, 28, 28, 1))
test_images <- test_images / 255

train_labels <- to_categorical(train_labels)
test_labels <- to_categorical(test_labels)

model %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)
              
model %>% fit(
  train_images, train_labels, 
  epochs = 5, batch_size=64
)
```

###Evaluating on test dataset

```{r, echo=TRUE, results='hide'}
results <- model %>% evaluate(test_images, test_labels)
results
```

```{r}
```




#Cats and dogs
##Organize downloaded images
```{r, echo=TRUE, results='hide'}
original_dataset_dir <- "~/Downloads/train"

base_dir <- "~/Downloads/cats_and_dogs_small"
dir.create(base_dir)

train_dir <- file.path(base_dir, "train")
dir.create(train_dir)
validation_dir <- file.path(base_dir, "validation")
dir.create(validation_dir)
test_dir <- file.path(base_dir, "test")
dir.create(test_dir)

train_cats_dir <- file.path(train_dir, "cats")
dir.create(train_cats_dir)

train_dogs_dir <- file.path(train_dir, "dogs")
dir.create(train_dogs_dir)

validation_cats_dir <- file.path(validation_dir, "cats")
dir.create(validation_cats_dir)

validation_dogs_dir <- file.path(validation_dir, "dogs")
dir.create(validation_dogs_dir)

test_cats_dir <- file.path(test_dir, "cats")
dir.create(test_cats_dir)

test_dogs_dir <- file.path(test_dir, "dogs")
dir.create(test_dogs_dir)

fnames <- paste0("cat.", 1:1000, ".jpg")
file.copy(file.path(original_dataset_dir, fnames), 
          file.path(train_cats_dir)) 

fnames <- paste0("cat.", 1001:1500, ".jpg")
file.copy(file.path(original_dataset_dir, fnames), 
          file.path(validation_cats_dir))

fnames <- paste0("cat.", 1501:2000, ".jpg")
file.copy(file.path(original_dataset_dir, fnames),
          file.path(test_cats_dir))

fnames <- paste0("dog.", 1:1000, ".jpg")
file.copy(file.path(original_dataset_dir, fnames),
          file.path(train_dogs_dir))

fnames <- paste0("dog.", 1001:1500, ".jpg")
file.copy(file.path(original_dataset_dir, fnames),
          file.path(validation_dogs_dir)) 

fnames <- paste0("dog.", 1501:2000, ".jpg")
file.copy(file.path(original_dataset_dir, fnames),
          file.path(test_dogs_dir))
```





##Convnet model
```{r}
library(keras)

model <- keras_model_sequential() %>% 
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(150, 150, 3)) %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_flatten() %>% 
  layer_dense(units = 512, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

summary(model)
```
##Compiling model
```{r}
model %>% compile(
  loss = "binary_crossentropy",
  optimizer = optimizer_rmsprop(lr = 1e-4),
  metrics = c("acc")
)
```
## Data preprocessing
```{r}
# All images will be rescaled by 1/255
train_datagen <- image_data_generator(rescale = 1/255)
validation_datagen <- image_data_generator(rescale = 1/255)

train_generator <- flow_images_from_directory(
  # This is the target directory
  train_dir,
  # This is the data generator
  train_datagen,
  # All images will be resized to 150x150
  target_size = c(150, 150),
  batch_size = 20,
  # Since we use binary_crossentropy loss, we need binary labels
  class_mode = "binary"
)

validation_generator <- flow_images_from_directory(
  validation_dir,
  validation_datagen,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "binary"
)
```

##Testing image generator
```{r}
batch <- generator_next(train_generator)
str(batch)

firstTrain<-batch[[1]][1,,,]
plot(as.raster(firstTrain,max=1))
```
##Training the model
```{r, echo=TRUE, results='hide'}
history <- model %>% fit_generator(
  train_generator,
  steps_per_epoch = 100,
  epochs = 5,
  validation_data = validation_generator,
  validation_steps = 50
)
```

It is good practice to always save your models after training:

```{r}
model %>% save_model_hdf5("cats_and_dogs_small_1.h5")
```

#Bird Images
```{r}
nikoDir<-"/mnt/raid/Niko"
imageDir<-file.path(nikoDir, "images")
dir.create(imageDir)

bluejayDir<-file.path(imageDir,"Bluejay")
bluejayImages<-list.files(bluejayDir)
length(bluejayImages)
bluejayTrain<-file.path(bluejayDir,"train")
dir.create(bluejayTrain)
bluejayValidate<-file.path(bluejayDir,"validate")
dir.create(bluejayValidate)
bluejayTest<-file.path(bluejayDir,"test")
dir.create(bluejayTest)

goldfinchDir<-file.path(imageDir,"Goldfinch")
goldfinchImages<-list.files(goldfinchDir)
length(goldfinchImages)
goldfinchTrain<-file.path(goldfinchDir,"train")
dir.create(goldfinchTrain)
goldfinchValidate<-file.path(goldfinchDir,"validate")
dir.create(goldfinchValidate)
goldfinchTest<-file.path(goldfinchDir,"test")
dir.create(goldfinchTest)
file.copy(file.path(goldfinchDir,goldfinchImages[1:45]),file.path(goldfinchTrain))
file.copy(file.path(goldfinchDir,goldfinchImages[46:60]),file.path(goldfinchValidate))
file.copy(file.path(goldfinchDir,goldfinchImages[61:75]),file.path(goldfinchTest))

cardinalDir<-file.path(imageDir,"Cardinal")
cardinalImages<-list.files(cardinalDir)
length(cardinalImages)
cardinalTrain<-file.path(cardinalDir,"train")
dir.create(cardinalTrain)
cardinalValidate<-file.path(cardinalDir,"validate")
dir.create(cardinalValidate)
cardinalTest<-file.path(cardinalDir,"test")
dir.create(cardinalTest)
file.copy(file.path(cardinalDir,cardinalImages[1:38]),file.path(cardinalTrain))
file.copy(file.path(cardinalDir,cardinalImages[39:53]),file.path(cardinalValidate))
file.copy(file.path(cardinalDir,cardinalImages[54:68]),file.path(cardinalTest))




```

#Building model
```{r}
model <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(28, 28, 1)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu")


```

